How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-28 tse-paper/content.tex
Review included incl page 3

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the analysis of system
logs which are produced during runtime of the system.
The first difference is that build logs are produced during runtime
of the build of the system, not during runtime of the system itself.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} shows
example lines from a system log.

The first goal in analyzing system log files is generally to
separate constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters.
Then, a log parser works on this intermediate representation.
Its output is an ordered list of timed events and their corresponding
parameter values~\cite{he2016evaluation}.
This structured log can serve as input to various machine learning and
data mining processes.

Some of the techniques developed to parse system logs can operate
on build logs as well, such as methods to retrieve the
values of variable parts in a log message, e.g.\ by using regular
expressions~\cite{nagappan2010abstracting,xu2009detecting}.
% This is the easiest example matching here
% amar et al are not in our literature survey
However, almost all of the techniques developed to analyze system logs
leverage their inherent event structure.
Build logs generally lack this structure and thus require different
bespoke approaches.
% if included: this explanation is not understandable.
% It needs to better embedded in the surrounding text
% One example is comparing execution traces to reference
% traces of intended behavior to detect anomalies.
% Amar et
% al.~\cite{amar2019mining} employed a similar approach to detect
% relevant lines in build logs.
% In this article, we focus on extracting a single specified information
% from the build log as a whole with chunk retrieval techniques.
% Chunk
% retrieval techniques are used as a part of log parsing

\subsection{Study Selection}
\Cref{fig:lit-survey} presents an overview of the study selection
process.
Since the field of build log analysis is relatively young, we
surveyed as broad a population of scientific material as possible.
We did not limit the sources to specific Software Engineering
venues~\cite{petersen2015guidelines}, but survey all sources
monitored by Google Scholar, including Bachelor's theses, Master's
theses, and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases because it is the search engine for academic articles with
the broadest and most current index, including preprints.

An initial Google Scholar search for variations of ``build log''
returned more than 2.2 million results (2020-03-15), many of which stem
from unrelated fields such as construction and wood working.
We refined the search criteria to exclude such obvious
non-related fields and also to exclude works on systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results
(2020-03-15).
% Our replication package documents the full search
% queries~\cite{brandt2020chunk-replication}.
To make handling so many search results feasible, we employed a
three-pass filtering strategy:

First, we filtered articles based on (a) their title and (b) the
contextual information Google Scholar displayed.
We included works at large that seemed to bear a resemblance to software
engineering and building software.
We excluded works written in a language unintelligible to the authors
(i.e., not English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

@@

Subsequently, we performed a second, finer filter phase by downloading the
full text of the works, reading their abstracts, and searching the
full text
for the occurrence of ``log.''
If a work showed traces of
working with build logs, we included it in the survey.
In total, we
were left with 87 works after the second filtering phases.

\subsection{Article Classification}
Having trimmed down the set of articles by 95\%, we investigated the
remaining ones in depth.
We wanted to characterize how research uses build logs.

Specifically, we were interested in
\begin{itemize}
  \item what information is retrieved from the logs and
  what the retrieved information is used for (\textbf{RQ1.1})
  \item which techniques are used to process the build logs,
  in how much detail they are described, and whether they are published
  as tools (\textbf{RQ1.2})
  \item what kind of build logs are used and the origin of these
  logs.
(\textbf{RQ1.3})
\end{itemize}

This lead us to the following research questions:
\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What is the state-of-the-research on analyzing build logs?}
% M: All of these are passive
% C: it says resolved, but they still are?
% I exchanged 'extracted' with targeted / analysis
% 'information retrieved' or 'information extracted' will give us
% this is not IR neither IE discussions again
\begin{itemize}[leftmargin=1cm]
  \item[\textbf{RQ1.1:}] What information is targeted in build logs?
  \item[\textbf{RQ1.2:}] Which analysis techniques exist?
  \item[\textbf{RQ1.3:}] Which kinds of build logs are subject to
  analysis?
\end{itemize}
\end{simplebox}

Since we had been careful to include rather than exclude borderline
works, this deeper evaluation excluded another 15 articles from the 87
works left after the second filter phase.
Multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the analyzing of build logs.
This excluded another 11 articles.
In sum, we extracted data from 61 articles.

To guide the data extraction we created a template.
The template form contained questions to answer RQ 1.1 through RQ 1.3,
for example ``Is the technique explained in detail?'' and
``What is the source of the build logs?''
As recommended by Petersen et
al.~\cite{petersen2015guidelines}, we started the mapping study with a
pilot study:
both authors extracted data from the same five articles and held a
consensus meeting to unify their understanding
of the data extraction template.
We divided the remaining 56 articles evenly between the two authors
and inspected each as closely as necessary to fill out the template
with confidence,
starting from occurrences of ``log'' or ``build'' within the article's
text.

For questions that did not require a boolean answer, we allowed assigning
multiple tags or categories as the first part (``preparation phase'')
of a virtual open card sorting~\cite{zimmermann2016card}.
An example for this is the source of the analyzed build logs,
where we assigned tags such as ``Travis CI'',
``Industrial'', ``Google'', or
``self built''.
Once we had classified all works, we went into the second phase
(``execution phase'') in which we grouped tags
together and found synonyms.
% the following adds to the example but is maybe unneeded / too much
For instance, in the case of the build log source we grouped
``Travis'' and ``travis ci'' together and assigned ``Industrial'' to
all articles that used build logs provided by a company.
The full data extraction template and our results of this step can be
found in our replication package~\cite{brandt2020chunk-replication}.

\addtolength{\tabcolsep}{-5pt}
\begin{table*}[tbhp]
\tinyish
\centering
\caption{Overview of build log analysis techniques.}
\begin{tabularx}{\textwidth}{@{}lXl@{}}

@@

\subsection{Results}
In this section, we present the results of the analysis phase of the
literature mapping along RQ 1.1 through RQ 1.3.

\subsubsection{What information is targeted in build logs? (RQ 1.1)}

The precise information targeted in the build logs varied from article
to article.
\Cref{fig:litsur:info_target} gives an overview over the frequency of
the targeted information.
Most prominent was the search for errors (39\%), followed by extracting
the executed or the failing
tests (both 11\%).
There were also numerous works which targeted more specialized
information, such as the environments used during the
build~\cite{zolfagharinia2017not}, packages
installed~\cite{selberg2012use}, or hints to source files which
relate to build failures~\cite{ren2018automated}.

>>>>

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the analysis of system
logs which are produced during runtime of the system.
The first difference is that build logs are produced during runtime
of the build of the system, not during runtime of the system itself.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} depicts two
examplary lines from such a system log.

The first goal in analyzing system log files is generally to
separate the constant and variable parts within a log
message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log
messages are clustered into log events, unifying messages with
identical constant parts and varying parameters.
Then, a log parser works on this intermediate representation.
Its output is an ordered list of timed events and their corresponding
parameter values~\cite{he2016evaluation}.
This structured log can serve as input to various machine learning and
data mining processes.

Some of the techniques developed to parse system logs can operate
on build logs as well, such as methods to retrieve the
values of variable parts in a log message, e.g.\, by using regular
expressions~\cite{nagappan2010abstracting,xu2009detecting}.
% This is the easiest example matching here
% amar et al are not in our literature survey
However, almost all of the techniques developed to analyze system logs
leverage their inherent event structure.
Build logs generally lack this structure and thus require different
bespoke approaches.
% if included: this explanation is not understandable.
% It needs to better embedded in the surrounding text
% One example is comparing execution traces to reference
% traces of intended behavior to detect anomalies.
% Amar et
% al.~\cite{amar2019mining} employed a similar approach to detect
% relevant lines in build logs.
% In this article, we focus on extracting a single specified information
% from the build log as a whole with chunk retrieval techniques.
% Chunk
% retrieval techniques are used as a part of log parsing

\subsection{Literature Selection}
The first step in a systematic literature study is to establish which works one wants to cover and how to find them~\cite{kitchenham2009systematic}.
Since the field of build log analysis is relatively young, we
surveyed as broad a population of scientific material as possible.
We did not limit the sources to specific Software Engineering
venues~\cite{petersen2015guidelines}, but included all sources
monitored by Google Scholar, including Bachelor's theses, Master's
theses, and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases because it is the search engine for scholarly material with
the broadest and most current index, including preprints.

\Cref{fig:lit-survey} gives an overview of the study selection
process. An initial Google Scholar search for variations of ``build log''
returned more than 2.2 million results on the 15th of March 2020, many of which stem
from unrelated fields such as construction or wood working.
We refined the search criteria to exclude such obvious
non-related fields and to exclude works on systems
or event logs (see \Cref{sec:system-log-analysis}).
This left us with 1,552 search results.
% Our replication package documents the full search
% queries~\cite{brandt2020chunk-replication}.
To make handling so many search results feasible, we employed a
three-pass filtering strategy:

First, we filtered articles based on (a) their title and (b) the
contextual information Google Scholar displayed on its search results overview page.
We included works at large that seemed to bear a resemblance to software
engineering and building software.
Following Kitchham's protocol~\cite{kitchenham2009systematic}, we excluded works written in a language unintelligible to the authors
(i.e., not English or German), which disregarded fewer than 1\% of search
hits.
We then unified
the results of the four search queries based on the link as the
identifying element, removing 60 articles that appeared in more than
one search query.
We removed further duplicates based on the title and were
left with 256 (16\%) of the original search results.
% removed duplicates of data extraction pass already here

@@

Second, we performed a finer filter phase by (1) reading their abstracts, and (2) downloading the
full text of the works and searching the
full text
for the occurrence of ``log.''
If a work showed traces of
working with build logs, we included it for the next filter phase.
In total, we
were left with 87 works after the second filtering phases.

Third, since we had been careful to include rather than exclude borderline
works, a deeper evaluation excluded another 15 articles from the 87
works left after the second filter phase.
Multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the analyzing of build logs.
This excluded another 11 articles.
In sum, we extracted data from 61 articles.

\subsection{Literature Classification}
Having trimmed down the set of articles by 95\%, we investigated the
remaining ones in-depth.
We wanted to characterize how research works with build logs.

% TODO  everything is in passive in the following
Specifically, we were interested in
\begin{itemize}
  \item what information is retrieved from the logs and
  what the retrieved information is used for (\textbf{RQ1.1})
  \item which techniques are used to process the build logs,
  in how much detail they are described, and whether they are published
  as tools (\textbf{RQ1.2})
  \item what kind of build logs are used and the origin of these
  logs.
(\textbf{RQ1.3})
\end{itemize}

This lead us to the following research questions:
\begin{simplebox}[attach boxed title to top center={yshift=-6mm}]
{\textbf{RQ1:} What is the state-of-the-research on analyzing build logs?}
% M: All of these are passive
% C: it says resolved, but they still are?
% M: Only the first one still is (don't know how to convert it), the others
% are not?
% I exchanged 'extracted' with targeted / analysis
% 'information retrieved' or 'information extracted' will give us
% this is not IR neither IE discussions again
\begin{itemize}[leftmargin=1cm]
  \item[\textbf{RQ1.1:}] What information is targeted in build logs?
  \item[\textbf{RQ1.2:}] Which analysis techniques exist?
  \item[\textbf{RQ1.3:}] Which kinds of build logs are subject to
  analysis?
\end{itemize}
\end{simplebox}

To support the data extraction we created a template that we filled out for each of the 61 works.
The template contained questions to answer RQ 1.1 through RQ 1.3,
for example ``Is the technique explained in detail?'' and
``What is the source of the build logs?''
As recommended by Petersen et
al.~\cite{petersen2015guidelines}, we started the mapping study with a
pilot:
both authors extracted data from the same five articles and held a
consensus meeting to unify their understanding
of the data extraction template.
We divided the remaining 56 articles evenly between the two authors
and inspected each articlex as closely as necessary to fill out the template
with confidence,
starting from occurrences of ``log'' or ``build'' within the article's
text.

For questions that did not require a boolean answer, we allowed assigning
multiple tags or categories as the first part (``preparation phase'')
of a virtual open card sorting~\cite{zimmermann2016card}.
An example for this is the question ``What is the source of the build log?'',
where we assigned tags such as ``Travis CI,''
``Industrial,'' ``Google,'' or
``self-built.''
Once we had classified all works, we went into the second phase of card sorting
(``execution phase''), in which we grouped tags
together and found synonyms.
% the following adds to the example but is maybe unneeded / too much
For instance, in the case of the build log source we grouped
``Travis'' and ``Travis CI'' together and assigned ``Industrial'' to
all articles that used build logs provided by a company.
The replication package contains the data extraction template and all results of this step~\cite{brandt2020chunk-replication}.

\subsection{Descriptive Study Statistics}
To be done.

\subsection{Results}
\addtolength{\tabcolsep}{-5pt}
\begin{table*}[tbhp]
\tinyish
\centering
\caption{Overview of build log analysis techniques.}
\begin{tabularx}{\textwidth}{@{}lXl@{}}

@@

In this section, we present the results of the analysis phase of the
literature mapping along RQ 1.1 through RQ 1.3.

\subsubsection{What information is targeted in build logs? (RQ 1.1)}
The precise information targeted in the build logs varied from article
to article.
\Cref{fig:litsur:info_target} gives an overview over the frequency of
the targeted information.
Most prominent was the search for errors (39\%), followed by extracting
the executed or the failing
tests (both 11\%).
There were also numerous works which targeted more specialized
information, such as the environments used during the
build~\cite{zolfagharinia2017not}, packages
installed~\cite{selberg2012use}, or hints to source files which
relate to build failures~\cite{ren2018automated}.

### 2020-04-28 tse-paper/paper.tex
new study explanation draft
Fix BibTeX
Review included incl page 3

\newcommand{\smallish}{\fontsize{6.15pt}{20pt}\selectfont}
\newcommand{\tinyish}{\fontsize{6.5pt}{6.15pt}\selectfont}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Build logs are the textual by-products that automatic software build processes
create, often as part of their Continuous Integration (CI)
pipeline.
Build logs can serve as a paramount source of
information for developers to understand and debug depdency, compilation, or
test failures.
Manually
extracting the important chunks of information in a build log, though, is akin to
finding a needle in a haystack.
Recently, practitioners and researchers have started attempts to partly
automate
this time-consuming activity, to not only ease
developers' task of finding this needle but also to enable a novel class of
automated on-ward processing applications.
In this paper, we give a systematic overview of the emerging field of build log analysis.
In a large-scale literature mapping study, we first survey and categorize the existing methods to
extract information chunks from build logs.
In the second part of the paper, we develop prototypical implementations of the three most promising
techniques we found, namely program synthesis
by example, textual similarity, and keyword search.
Finally, we evaluate the techniques in an empirical study on
almost 800 build logs 
from a wide range of different projects and programming languages.
Our findings show that none of the three techniques in general outperforms
the others, but rather, that each technique has its respective strengths
and weaknesses.
We discuss under which circumstances each technique performs best
and provide a recommendation scheme for when developers or researchers might 
use which technique.
We close with a roadmap of important open problems in the
emerging field of
build log analysis.
\end{abstract}

>>>>

\newcommand{\smallish}{\fontsize{6.15pt}{20pt}\selectfont}
\newcommand{\tinyish}{\fontsize{6.5pt}{6.15pt}\selectfont}
\newcommand{\circlenum}[1]{\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {#1}}}}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.

% TODO: No spaces before/after ---
\title{How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Build logs are the textual by-products that automatic software build processes
create, often as part of their Continuous Integration (CI)
pipeline.
Build logs can serve as a paramount source of
information for developers to understand and debug depdency, compilation, or
test failures.
Manually
extracting the important chunks of information in a build log, though, is akin to
finding a needle in a haystack.
Recently, practitioners and researchers have started attempts to partly
automate
this time-consuming activity, to not only ease
developers' task of finding this needle but also to enable a novel class of
automated on-ward processing applications.
In this paper, we give a systematic overview of the emerging field of build log analysis.
In a large-scale literature mapping study, we first survey and categorize the existing methods to
extract information chunks from build logs.
In the second part of the paper, we develop prototypical implementations of the three most promising
techniques we found, namely program synthesis
by example, textual similarity, and keyword search.
Finally, we evaluate the techniques in an empirical study on
almost 800 build logs 
from a wide range of different projects and programming languages.
Our findings show that none of the three techniques in general outperforms
the others, but rather, that each technique has its respective strengths
and weaknesses.
We discuss under which circumstances each technique performs best
and provide a recommendation scheme for when developers or researchers might 
use which technique.
We close with a roadmap of important open problems in the field of
build log analysis.
\end{abstract}
